On particular examples of planar integral point sets and their classification \footnote{ This work was carried out at Voronezh State University and supported by the Russian Science Foundation grant 19-11-00197. }

### 2021-02-24 AvdMomZvolA_examples/ips_examples.tex
[AMZAE] Remove old intro

\begin{dfn}
	The function $d(m, n)$ is the minimum possible diameter of
	the integral point set $\mathcal{P}$ of $n$ points in
	$m$-dimensional Euclidean space $\mathbb{R}^{m}$.
\end{dfn}

%For a bit more sophisticated configurations, another functions are presented~\cite{kurz2008minimum,kreisel2008heptagon}.
%
%\begin{dfn}
%	The function $\overline{d}(m, n)$ is the minimum possible diameter of
%	the integral point set $\mathcal{P}$ of $n$ points in
%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
%	such that no $m+1$ points are located on an $(m-1)$-dimensional hyperplane.
%\end{dfn}

%\begin{dfn}
%	The function $\dot{d}(m, n)$ is the minimum possible diameter of
%	the integral point set $\mathcal{P}$ of $n$ points in
%	$m$-dimensional Euclidean space $\mathbb{R}^{m}$
%	such that no three points are located on a straight line
%	and no four points are located on a circle.
%\end{dfn}

For the list of known bounds for $d(m, n)$,
we refer the reader to~\cite[Theorem 1]{kurz2008bounds} or to~\cite{our-vmmsh-2018}.
We will discuss the following estimations presented at~\cite{kurz2008bounds}:
\begin{align}
	d(m, 2m + 1) \leq 8\\
	d(m, 2m + 2) \leq 13 \hypertarget{d2}\\
	d(m, 3m) \leq 109
\end{align}
and the following theorem \cite[Theorem 2.1]{kurz2008bounds}.

\begin{theorem}
	Let $\mathcal{P}$ be a planar integral point set consisting of
	$n - 2$ points on line $l_1$ and two points $P_{1}$ and $P_{2}$ on a
	parallel line $l_2$ with distance $r$ between $l_{1}$ and $l_{2}$. If there
	exist positive integers $v$, $w$ with $f^{2} + v^{2}
	= w^{2}$ and $v < 2r$, where $|P_{1}P_{2}| = f$,
	then

	\begin{equation}\label{formula1}
		d(m, n + 2(m - 2)) \leq \max(w, \operatorname{diam(\mathcal{P})})
	\end{equation}

\end{theorem}

Firstly, we discuss the classification of planar integral points sets;
then, we present some bounds for $d(m,n)$ based on planar integral point sets of particular types
and provide some general constructions for such bounds.

\section{Classification of planar integral point sets}

\subsection{Integral point sets situated in two straight lines}

\begin{dfn}
	A planar integral point sets of $n$ points with $n-1$ points on a straight line is called
	a \textit{facher} set.
\end{dfn}
Facher sets are very dominating examples of planar integral pont sets.
In~\cite{antonov2008maximal}, facher sets of characteristic 1 are called \textit{semi-crabs}.
For $9 \leq n \leq 122$, the diameter $d(2,n)$ is reached on a facher point set~\cite{kurz2008minimum}.

For non-facher integral point sets situated in two straight lines,
we can easily distinct the following three cases:

\begin{dfn}
	A planar integral point sets situated in two parallel straight lines
	is called a \textit{rails} set.
\end{dfn}

Among the rails sets, sets with 2 points on one line and all the other on another line dominate.

\begin{dfn}
	A planar integral point sets situated in two perpendicular straight lines
	is called a \textit{cross} set.
\end{dfn}
Every cross set has characteristic 1;
in~\cite{antonov2008maximal}, cross sets with only 2 points out of one of the lines are called \textit{crabs}.

\begin{dfn}
	A planar integral point sets situated in two straight lines
	that are not parallel nor perpendicular,
	is called a \textit{sciccors} set.
\end{dfn}

There is an important subclassof scissors sets.

\begin{dfn}
	A scissors set with an axis of symmetry,
	which is the angle bisector for the straight lines,
	is called a \textit{pyramid} set.
\end{dfn}

%TODO: rephrase?

%TODO: can a scissors set have another axis of symmetry?

\subsection{Other integral point sets}

\begin{dfn}
	A planar integral point sets that is situated on a circle is called a \textit{circular}
	point set.
\end{dfn}

Circular sets are very important examples
of integral point sets~\cite{harborth1993upper,piepmeyer1996maximum,bat2018number}.

\begin{dfn}
	A planar integral point sets that is situated on the conjunction of a circle with its center,
	is called a \textit{centered-circular} point set.
\end{dfn}

These six classes dominate among all the known planar integral point sets;
however, some sophisticated constructions are also known.

%TODO: pictures!

>>>>

While the minimal possible diameter for planar integral point sets of given cardinality was being counted,
it was noticed~\cite{kurz2008minimum} that such diameter is attained at sets with many points on a straight line;
for some estimations on this tendency, we also refer the reader to~\cite{solymosi2003note}.
